---
title: "Opioid-Related Death in Massachusetts"
subtitle: "Simulated Demo"
author: "Gabby Novak"
output: pdf_document
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=F,
                      error=F,
                      warning=F,
                      fig.height=11,
                      fig.width=8)
setwd("D:/Simulated-Research-Demos")
```

```{r, results="hide"}
#### Packages ####
library(tidyverse)
# Loads: ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, & forcats
library(tigris)
library(lubridate)
library(knitr)
library(kableExtra)
library(grid)
library(gridExtra)
library(extrafont)
library(viridis)

#### External Data ####
# base shp file for mapping
shp<-zctas(year=2010,state="Massachusetts")
# Example occupations and industries
occups<-read_csv("Occups.csv")
# ICD10 2018 codes
# Source: https://www.cms.gov/Medicare/Coding/ICD10/2018-ICD-10-CM-and-GEMs.html
icd10<-read_csv("icd10cm_order_2018.csv")
```

# Data Set Unification

This project made use of publicly available death records of individuals who died in Massachusetts, USA between 2000 and 2017 with an opioid-related ICD10 code assigned to them as a cause of death. The data source presented several challenges, not least among them, errors due to manual entry of information. However, the greatest hurdle were changes to data format in mid-2014. Prior to that 2014 format change, 77 variables were available. After the format change, a staggering 843 were available. While the vast majority of the earlier variables were repeated in the newer format, both variable names and coding structures were updated. In order to fit any sort of temporal model, we needed a unified data set. Additionally, while we were interested in several individual-level covariates, much of these data were irrelevant to us. I developed the following functions as a mechanism to extract specific information from the raw vitals, recode it, and populate a data frame much more suited to the project's needs. Note that these functions create coded data sets. The full data included several hundred thousand observations. The coding was a mechanism meant to reduce file size for sharing between colleagues. The analysis was actually completed using data run through another function which converted the numeric labels to their representative values. 

```{r}
#### For deaths 2000-mid 2014 ####
vital.00.14<-function(dataset){
  temp<-list()
  state<-c("ALABAMA","ALASKA","ARIZONA","ARKANSAS","CALIFORNIA","COLORADO",
           "CONNECTICUT","DELAWARE","FLORIDA","GEORGIA","HAWAII","IDAHO",
           "ILLINOIS","INDIANA","IOWA","KANSAS","KENTUCKY","LOUISIANA",
           "MAINE","MARYLAND","MASSACHUSETTS","MICHIGAN","MINNESOTA",
           "MISSISSIPPI","MISSOURI","MONTANA","NEBRASKA","NEVADA",
           "NEW HAMPSHIRE","NEW JERSEY","NEW MEXICO","NEW YORK",
           "NORTH CAROLINA","NORTH DAKOTA","OHIO","OKLAHOMA","OREGON",
           "PENNSYLVANIA","RHODE ISLAND","SOUTH CAROLINA","SOUTH DAKOTA",
           "TENNESSEE","TEXAS","UTAH","VERMONT","VIRGINIA","WASHINGTON",
           "WASHINGTON DC","WEST VIRGINIA","WISCONSIN","WYOMING")
  abbr<-c("AL","AK","AZ","AR","CA","CO","CT","DC","DE","FL","GA","HI","ID",
          "IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO",
          "MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI",
          "SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY")
  # batch
  temp$batch<-1
  for(j in 1:nrow(dataset)){
    # sfnum
    temp$sfnum[j]<-unlist(dataset[j,"CERT"])
    # ddate
    temp$ddate[j]<-paste0(str_sub(dataset[j,"DOD"],1,4),"-",
                                  str_sub(dataset[j,"DOD"],5,6),"-",
                                  str_sub(dataset[j,"DOD"],7,8))
    # male
    if(dataset[j,"SEX"]=="1"){temp$male[j]<-1}
    if(dataset[j,"SEX"]=="2"){temp$male[j]<-0}
    # age
    if(str_sub(dataset[j,"AGE_AT_DEATH"],1,1)==0|
       str_sub(dataset[j,"AGE_AT_DEATH"],1,1)==1)
    {temp$age[j]<-as.numeric(str_sub(dataset[j,"AGE_AT_DEATH"],-2,-1))}
    else{if(str_sub(dataset[j,"AGE_AT_DEATH"],1,1)==2|
            str_sub(dataset[j,"AGE_AT_DEATH"],1,1)==4|
            str_sub(dataset[j,"AGE_AT_DEATH"],1,1)==5|
            str_sub(dataset[j,"AGE_AT_DEATH"],1,1)==6)
      {temp$age[j]<-0}else{if(str_sub(dataset[j,"AGE_AT_DEATH"],1,1)==9){temp$age[j]<-NA}}}
    # race
    if(!(dataset[j,"DETHNIC_HISPANIC"]=="0"|dataset[j,"DETHNIC_HISPANIC"]=="9"))
      {temp$race[j]<-3}
      else{if(dataset[j,"RACE"]=="01"){temp$race[j]<-1}
           if(dataset[j,"RACE"]=="02"){temp$race[j]<-2}
           if(dataset[j,"RACE"]=="03"){temp$race[j]<-5}
           if(dataset[j,"RACE"]=="04"|
              dataset[j,"RACE"]=="05"|
              dataset[j,"RACE"]=="06"|
              dataset[j,"RACE"]=="07"|
              dataset[j,"RACE"]=="08"|
              dataset[j,"RACE"]=="09"|
              dataset[j,"RACE"]=="10"|
              dataset[j,"RACE"]=="11"|
              dataset[j,"RACE"]=="12"){temp$race[j]<-4}
           if(dataset[j,"RACE"]=="13"|
              dataset[j,"RACE"]=="14"){temp$race[j]<-7}
           if(dataset[j,"RACE"]=="99"){temp$race[j]<-NA}}
    # occup
    temp$occup[j]<-unlist(dataset[j,"OCCUP"])
    # indust
    temp$indust[j]<-unlist(dataset[j,"INDUST"])
    # edu
    if(as.numeric(dataset[j,"DEDUC"])<=11){temp$edu[j]<-1}
      else{if(as.numeric(dataset[j,"DEDUC"])<=13){temp$edu[j]<-2}
        else{if(as.numeric(dataset[j,"DEDUC"])<=16){temp$edu[j]<-3}
          else{if(dataset[j,"DEDUC"]=="99"){temp$edu[j]<-NA}
            else{if(as.numeric(dataset[j,"DEDUC"])>16){temp$edu[j]<-4}}}}}
    # immig
    if(dataset[j,"NATIVITY"]=="99"){temp$immig[j]<-NA}
      else{if(as.numeric(dataset[j,"NATIVITY"])>51){temp$immig[j]<-4}
        else{temp$immig[j]<-5}}
    # pimmig
    ifelse(!dataset[j,"FATHER_BSTATE"]%in%state&
           !dataset[j,"FATHER_BSTATE"]%in%abbr&
           !dataset[j,"FATHER_BSTATE"]=="UNKNOWN",
           yes=ifelse(!dataset[j,"MOTHER_BSTATE"]%in%state&
                      !dataset[j,"MOTHER_BSTATE"]%in%abbr&
                      !dataset[j,"MOTHER_BSTATE"]=="UNKNOWN",
                      yes=temp$pimmig[j]<-2,
                      no=temp$pimmig[j]<-1),
           no=ifelse(!dataset[j,"MOTHER_BSTATE"]%in%state&
                     !dataset[j,"MOTHER_BSTATE"]%in%abbr&
                     !dataset[j,"MOTHER_BSTATE"]=="UNKNOWN",
                     yes=temp$pimmig[j]<-1,
                     no=ifelse((dataset[j,"FATHER_BSTATE"]%in%state|
                         dataset[j,"FATHER_BSTATE"]%in%abbr)&
                        (dataset[j,"MOTHER_BSTATE"]%in%state|
                         dataset[j,"MOTHER_BSTATE"]%in%abbr),
                        yes=temp$pimmig[j]<-0,
                        no=temp$pimmig[j]<-NA)))
    # marital
    if(dataset[j,"MARITAL"]=="1"){temp$marital[j]<-5}
    if(dataset[j,"MARITAL"]=="2"){temp$marital[j]<-1}
    if(dataset[j,"MARITAL"]=="3"){temp$marital[j]<-3}
    if(dataset[j,"MARITAL"]=="4"){temp$marital[j]<-4}
    if(dataset[j,"MARITAL"]=="9"){temp$marital[j]<-NA}
    # veteran
    if(dataset[j,"VET_STAT"]==0){temp$veteran[j]<-0}
    else{if(dataset[j,"VET_STAT"]==9){temp$veteran[j]<-NA}
      else{temp$veteran[j]<-1}}
    # preg
    temp$preg[j]<-NA
    # resadd
    temp$resadd[j]<-str_remove_all(paste(dataset[j,"RES_ADDR_NUM"],
                                           dataset[j,"RES_ADDR1"],
                                           dataset[j,"RES_STREET_DESIG"])," NA")
    # rescity
    temp$rescity[j]<-unlist(dataset[j,"RES_CITY"])
    # resstate
    ifelse(is.na(dataset[j,"RES_CITY_CODE"]),
           yes=temp$resstate[j]<-NA,
           no=ifelse(as.numeric(dataset[j,"RES_CITY_CODE"])<=351,
                     yes=temp$resstate[j]<-"MASSACHUSETTS",
                     no=temp$resstate[j]<-"OUT OF STATE"))
    # reszip
    temp$reszip[j]<-unlist(dataset[j,"RES_ZIP"])
    # resnat
    temp$resnat[j]<-NA
    # dplace
    ifelse(dataset[j,"DPLACE"]==1,
           yes=temp$dplace[j]<-1,
           no=ifelse(dataset[j,"DPLACE"]==2,
                     yes=temp$dplace[j]<-2,
                     no=ifelse(dataset[j,"DPLACE"]==3,
                               yes=temp$dplace[j]<-3,
                               no=ifelse(dataset[j,"DPLACE"]==5,
                                         yes=temp$dplace[j]<-6,
                                         no=ifelse(dataset[j,"DPLACE"]==6,
                                                   yes=temp$dplace[j]<-4,
                                                   no=ifelse(dataset[j,"DPLACE"]==7,
                                                             yes=temp$dplace[j]<-8,
                                                             no=temp$dplace[j]<-NA))))))
    # dfacilitynum
    if(dataset[j,"FACCODE"]=="0000"|
       dataset[j,"FACCODE"]=="0060"|
       dataset[j,"FACCODE"]=="0070"|
       dataset[j,"FACCODE"]=="0080"|
       dataset[j,"FACCODE"]=="0090"|
       dataset[j,"FACCODE"]=="9999"){temp$dfacilitynum[j]<-NA}
    else{temp$dfacilitynum[j]<-unlist(dataset[j,"FACCODE"])}
    # dadd
    temp$ddad[j]<-NA
    # dcity
    temp$dcity[j]<-unlist(dataset[j,"DNAME_CITY"])
    # dstate
    ifelse(dataset[j,"DSTATEL"]=="MA",
           yes=temp$dstate[j]<-"MASSACHUSETTS",
           no=ifelse(dataset[j,"DSTATEL"]=="MASSACHUSETTS",
                     yes=temp$dstate[j]<-"MASSACHUSETTS",
                     no=temp$dstate[j]<-NA))
    # dzip
    temp$dzip[j]<-NA
    # dnat
    temp$dnat[j]<-"UNITED STATES"
    # travel
    ifelse(!is.na(dataset[j,"RES_CITY"])&!is.na(dataset[j,"DNAME_CITY"]),
           yes=ifelse(dataset[j,"RES_CITY"]==dataset[j,"DNAME_CITY"],
                      yes=temp$travel[j]<-0,
                      no=temp$travel[j]<-1),
           no=temp$travel[j]<-NA)
    # All icd variables
    y<-str_trim(str_split(str_replace_all(dataset[j,"TRX_REC_AXIS_CD"],"  "," "),
                          " ",simplify=T),side="both")
    x<-vector(mode="character")
    l<-1
    for(k in 1:length(y)){
      if(str_length(y[k])>4){
        if(str_length(y[k])<6)
        {x[l]<-str_remove(y[k],".$")
        l<-l+1}
        else
        {x[l]<-str_split(y[k],"0",simplify=T)[1]
        l<-l+1
        x[l]<-str_split(y[k],"0",simplify=T)[2]
        l<-l+1}}
      else
      {x[l]<-y[k]
      l<-l+1}}
    # icd1
    if(is.na(x[1])){temp$icd1[j]<-NA}
    else{temp$icd1[j]<-x[1]}
    # icd2
    if(is.na(x[2])){temp$icd2[j]<-NA}
    else{temp$icd2[j]<-x[2]}
    # icd3
    if(is.na(x[3])){temp$icd3[j]<-NA}
    else{temp$icd3[j]<-x[3]}
    # icd4
    if(is.na(x[4])){temp$icd4[j]<-NA}
    else{temp$icd4[j]<-x[4]}
    # icd5
    if(is.na(x[5])){temp$icd5[j]<-NA}
    else{temp$icd5[j]<-x[5]}
    # icd6
    if(is.na(x[6])){temp$icd6[j]<-NA}
    else{temp$icd6[j]<-x[6]}
    # icd7
    if(is.na(x[7])){temp$icd7[j]<-NA}
    else{temp$icd7[j]<-x[7]}
    # icd8
    if(is.na(x[8])){temp$icd8[j]<-NA}
    else{temp$icd8[j]<-x[8]}
    # icd9
    if(is.na(x[9])){temp$icd9[j]<-NA}
    else{temp$icd9[j]<-x[9]}
    # icd10
    if(is.na(x[10])){temp$icd10[j]<-NA}
    else{temp$icd10[j]<-x[10]}
    # icd11
    if(is.na(x[11])){temp$icd11[j]<-NA}
    else{temp$icd11[j]<-x[11]}
    # icd12
    if(is.na(x[12])){temp$icd12[j]<-NA}
    else{temp$icd12[j]<-x[12]}
    # icd13
    if(is.na(x[13])){temp$icd13[j]<-NA}
    else{temp$icd13[j]<-x[13]}
    #icd14
    if(is.na(x[14])){temp$icd14[j]<-NA}
    else{temp$icd14[j]<-x[14]}
    # icd15
    if(is.na(x[15])){temp$icd15[j]<-NA}
    else{temp$icd15[j]<-x[15]}
    #icd16
    if(is.na(x[16])){temp$icd16[j]<-NA}
    else{temp$icd16[j]<-x[16]}
  }
  return(as_tibble(temp))
}
```

\newpage

```{r}
#### For deaths late 2014-2017 ####
vital.14.17<-function(dataset){
  temp<-list()
  # batch
  temp$batch<-2
  for(i in 1:nrow(dataset)){
    # sfnum
    temp$sfnum[i]<-unlist(dataset[i,"SFN_NUM"])
    # ddate
    temp$ddate[i]<-paste0(str_sub(dataset[i,"DOD_4_FD"],7,10),"-",
                          str_sub(dataset[i,"DOD_4_FD"],1,2),"-",
                          str_sub(dataset[i,"DOD_4_FD"],4,5))
    # male
    ifelse(dataset[i,"SEX"]=="M",
           yes=temp$male[i]<-1,
           no=ifelse(dataset[i,"SEX"]=="F",
                     yes=temp$male[i]<-0,
                     no=temp$male[i]<-NA))
    # age
    if(dataset[i,"AGETYPE"]==1)
    {temp$age[i]<-unlist(dataset[i,"AGE1_CALC"])}
    else{if(dataset[i,"AGETYPE"]==2|
            dataset[i,"AGETYPE"]==3)
    {temp$age[i]<-0}
      else{if(dataset[i,"AGETYPE"]==8|
              dataset[i,"AGETYPE"]==9)
      {temp$age[i]<-NA}}}
    # race
    ifelse(str_count(paste0(dataset[i,"RACE1"],
                            dataset[i,"RACE_AM_NATIVE"],
                            dataset[i,"RACE_ASIAN"],
                            dataset[i,"RACE_BLACK"],
                            dataset[i,"DETHNIC4"]),"Y")>1,
           yes=temp$race[i]<-6,
           no=ifelse(dataset[i,"RACE_HISP_LAT_WHITE"]=="Y"|
                     dataset[i,"RACE_HISP_LAT_BLACK"]=="Y"|
                     dataset[i,"DETHNIC4"]=="Y",
                     yes=temp$race[i]<-3,
                     no=ifelse(dataset[i,"RACE1"]=="Y",
                               yes=temp$race[i]<-1,
                               no=ifelse(dataset[i,"RACE_BLACK"]=="Y",
                                         yes=temp$race[i]<-2,
                                         no=ifelse(dataset[i,"RACE_ASIAN"]=="Y",
                                                   yes=temp$race[i]<-4,
                                                   no=ifelse(dataset[i,"RACE_AM_NATIVE"]=="Y",
                                                             yes=temp$race[i]<-5,
                                                             no=ifelse(dataset[i,"RACE_UNK"]=="Y",
                                                                       yes=temp$race[i]<-NA,
                                                                       no=temp$race[i]<-7)))))))
    # occup
    temp$occup[i]<-unlist(dataset[i,"OCCUP"])
    # indust
    temp$indust[i]<-unlist(dataset[i,"INDUST"])
    # edu
    ifelse(dataset[i,"DEDUC"]==1|
           dataset[i,"DEDUC"]==2,
           yes=temp$edu[i]<-1,
           no=ifelse(dataset[i,"DEDUC"]==3|
                     dataset[i,"DEDUC"]==4|
                     dataset[i,"DEDUC"]==5,
                     yes=temp$edu[i]<-2,
                     no=ifelse(dataset[i,"DEDUC"]==6|
                               dataset[i,"DEDUC"]==7,
                               yes=temp$edu[i]<-3,
                               no=ifelse(dataset[i,"DEDUC"]==8|
                                         dataset[i,"DEDUC"]==9,
                                         yes=temp$edu[i]<-4,
                                         no=ifelse(dataset[i,"DEDUC"]==12,
                                                   yes=temp$edu[i]<-5,
                                                   no=temp$edu[i]<-NA)))))
    # immig
    ifelse(dataset[i,"RES_COUNTRY"]=="UNITED STATES",
           yes=ifelse(dataset[i,"BPLACE_CNT"]=="UNITED STATES",
                      yes=temp$immig[i]<-0,
                      no=temp$immig[i]<-1),
           no=ifelse(dataset[i,"BPLACE_CNT"]=="UNITED STATES",
                     yes=temp$immig[i]<-3,
                     no=temp$immig[i]<-2))
    # pimmig
    ifelse(!(dataset[i,"FATHER_BCOUNTRY"]=="UNITED STATES"|
           dataset[i,"FATHER_BCOUNTRY"]=="UNKNOWN"),
           yes=ifelse(!(dataset[i,"MOTHER_BCOUNTRY"]=="UNITED STATES"|
                      dataset[i,"MOTHER_BCOUNTRY"]=="UNKNOWN"),
                      yes=temp$pimmig[i]<-2,
                      no=temp$pimmig[i]<-1),
           no=ifelse(!(dataset[i,"MOTHER_BCOUNTRY"]=="UNITED STATES"|
                     dataset[i,"MOTHER_BCOUNTRY"]=="UNKNOWN"),
                     yes=temp$pimmig[i]<-1,
                     no=ifelse(dataset[i,"FATHER_BCOUNTRY"]=="UNITED STATES"&
                               dataset[i,"MOTHER_BCOUNTRY"]=="UNITED STATES",
                               yes=temp$pimmig[i]<-0,
                               no=temp$pimmig[i]<-NA)))
    # marital
    ifelse(dataset[i,"MARITAL"]=="M"|
           dataset[i,"MARITAL"]=="A",
           yes=temp$marital[i]<-1,
           no=ifelse(dataset[i,"MARITAL"]=="W",
                     yes=temp$marital[i]<-3,
                     no=ifelse(dataset[i,"MARITAL"]=="D",
                               yes=temp$marital[i]<-4,
                               no=ifelse(dataset[i,"MARITAL"]=="S",
                                         yes=temp$marital[i]<-5,
                                         no=temp$marital[i]<-NA))))
    # veteran
    ifelse(dataset[i,"ARMED"]=="Y",
           yes=temp$veteran[i]<-1,
           no=ifelse(dataset[i,"ARMED"]=="N",
                     yes=temp$veteran[i]<-0,
                     no=temp$veteran[i]<-NA))
    # preg
    if(is.na(dataset[i,"PREG"])){temp$preg[i]<-NA}
    else{if(dataset[i,"PREG"]==1){temp$preg[i]<-0}
    else{if(dataset[i,"PREG"]==2){temp$preg[i]<-1}
         else{if(dataset[i,"PREG"]==3|dataset[i,"PREG"]==4){temp$preg[i]<-2}
           else{temp$preg[i]<-NA}}}}
  # resadd
  temp$resadd[i]<-str_remove_all(paste(dataset[i,"RES_ADDR_NUM"],
                                       dataset[i,"RES_STREET_PREFIX"],
                                       dataset[i,"RES_ADDR1"],
                                       dataset[i,"RES_STREET_DESIG"],
                                       dataset[i,"RES_STREET_SUFFIX"],
                                       dataset[i,"RES_ADDR2"]),
                                 " NA")
  # rescity
  temp$rescity[i]<-unlist(dataset[i,"RES_CITY"])
  # resstate
  temp$resstate[i]<-unlist(dataset[i,"RES_STATE"])
  # reszip
  temp$reszip[i]<-unlist(dataset[i,"RES_ZIP"])
  # resnat
  temp$resnat[i]<-unlist(dataset[i,"RES_COUNTRY"])
  # dplace
  ifelse(dataset[i,"DPLACE"]==9,
         yes=temp$dplace[i]<-NA,
         no=temp$dplace[i]<-unlist(dataset[i,"DPLACE"]))
  # dfacilitynum
  temp$dfacilitynum[i]<-unlist(dataset[i,"DFACILITYL"])
  # dadd
  temp$ddad[i]<-str_remove_all(paste(dataset[i,"DADDR_NUM"],
                                     dataset[i,"DSTREET_PREFIX"],
                                     dataset[i,"DADDR1"],
                                     dataset[i,"DSTREET_DESIG"],
                                     dataset[i,"DSTREET_SUFFIX"],
                                     dataset[i,"DADDR2"]),
                               " NA")
  # dcity
  temp$dcity[i]<-unlist(dataset[i,"DNAME_CITY"])
  # dstate
  temp$dstate[i]<-unlist(dataset[i,"DSTATEL"])
  # dzip
  temp$dzip[i]<-str_extract(dataset[i,"DZIP9"],"^.{5}")
  # dnat
  temp$dnat[i]<-unlist(dataset[i,"DCOUNTRY"])
  # travel
  ifelse(!is.na(dataset[i,"DNAME_CITY"])&
         !is.na(dataset[i,"RES_CITY"]),
         yes=ifelse(!dataset[i,"DNAME_CITY"]==dataset[i,"RES_CITY"],
                    yes=temp$travel[i]<-1,
                    no=temp$travel[i]<-0),
         no=temp$travel[i]<-NA)
  # icd1
  z<-str_trim(str_split(str_replace_all(dataset[i,"TRX_REC_AXIS_CD"],"  "," "),
                        " ",simplify=T),side="both")
  y<-vector(mode="character")
  l<-1
  for(k in 1:length(z)){
    if(str_length(z[k])>4){
      if(str_length(z[k])<6)
      {y[l]<-str_remove(z[k],".$")
      l<-l+1}
      else
      {y[l]<-str_split(z[k],"0",simplify=T)[1]
      l<-l+1
      y[l]<-str_split(z[k],"0",simplify=T)[2]
      l<-l+1}}
    else
    {y[l]<-z[k]
    l<-l+1}}
  if(is.na(y[1])){temp$icd1[i]<-NA}
  else{temp$icd1[i]<-y[1]}
  # icd2
  if(is.na(y[2])){temp$icd2[i]<-NA}
  else{temp$icd2[i]<-y[2]}
  # icd3
  if(is.na(y[3])){temp$icd3[i]<-NA}
  else{temp$icd3[i]<-y[3]}
  # icd4
  if(is.na(y[4])){temp$icd4[i]<-NA}
  else{temp$icd4[i]<-y[4]}
  # icd5
  if(is.na(y[5])){temp$icd5[i]<-NA}
  else{temp$icd5[i]<-y[5]}
  # icd6
  if(is.na(y[6])){temp$icd6[i]<-NA}
  else{temp$icd6[i]<-y[6]}
  # icd7
  if(is.na(y[7])){temp$icd7[i]<-NA}
  else{temp$icd7[i]<-y[7]}
  # icd8
  if(is.na(y[8])){temp$icd8[i]<-NA}
  else{temp$icd8[i]<-y[8]}
  # icd9
  if(is.na(y[9])){temp$icd9[i]<-NA}
  else{temp$icd9[i]<-y[9]}
  # icd10
  if(is.na(y[10])){temp$icd10[i]<-NA}
  else{temp$icd10[i]<-y[10]}
  # icd11
  if(is.na(y[11])){temp$icd11[i]<-NA}
  else{temp$icd11[i]<-y[11]}
  # icd12
  if(is.na(y[12])){temp$icd12[i]<-NA}
  else{temp$icd12[i]<-y[12]}
  # icd13
  if(is.na(y[13])){temp$icd13[i]<-NA}
  else{temp$icd13[i]<-y[13]}
  #icd14
  if(is.na(y[14])){temp$icd14[i]<-NA}
  else{temp$icd14[i]<-y[14]}
  # icd15
  if(is.na(y[15])){temp$icd15[i]<-NA}
  else{temp$icd15[i]<-y[15]}
  #icd16
  if(is.na(y[16])){temp$icd16[i]<-NA}
  else{temp$icd16[i]<-y[16]}
  }
  return(as_tibble(temp))
}
```

\newpage

# Simulated Data

## Individual Data Set

The following data set mimics that created by the above functions. 

```{r}
#### Preparing occupation data ####
occup<-occups%>%
  transmute(var=map2_chr(.x=occup,.y=indust,.f=~paste(.x,.y,sep=";")))
occup<-sample(occup[[1]],500,replace=T)
occup<-data.frame(occup)%>%
  separate(occup, into=c("occup","indust"),sep=";")

#### Subsetting icd10 codes ####
# To replicate repetition in actual data set
icd10<-sample_n(icd10[icd10$Header==1,],20000)

#### Coded data ####
# For reproducability
set.seed(8282019)

ind<-data.frame(
  batch=sample(c(1,2),500,replace=T),
  sfnum=sample(0:999999,500,replace=F),
  ddate=sample(seq(as.Date("2000-01-01"),as.Date("2017-12-31"),by="day"),500,replace=T),
  male=sample(0:1,500,replace=T),
  age=sample(0:100,500,replace=T),
  race=sample(1:7,500,replace=T),
  occup=occup$occup,
  indust=occup$indust,
  edu=sample(1:5,500,replace=T),
  immig=sample(0:4,500,replace=T),
  pimmig=sample(0:2,500,replace=T),
  marital=sample(c(1,3:5),500,replace=T),
  veteran=sample(0:1,500,replace=T),
  preg=sample(0:2,500,replace=T),
  resadd="1234 CIRCLE ST",
  rescity= "ANYTOWN",
  resstate= "MASSACHUSETTS",
  reszip=sample(shp@data$ZCTA5CE10,500,replace=T),
  resnat="UNITED STATES",
  dplace=sample(1:8,500,replace=T),
  dfacilitynum="000000",
  ddad="1234 SQUARE ST",
  dcity="ANYTOWN",
  dstate="MASSACHUSETTS",
  dzip=sample(shp@data$ZCTA5CE10,500,replace=T),
  dnat="UNITED STATES",
  travel=sample(0:1,500,replace=T),
  icd1=sample(icd10$ICD10Code,500,replace=T),
  icd2=sample(icd10$ICD10Code,500,replace=T),
  icd3=sample(icd10$ICD10Code,500,replace=T),
  icd4=sample(icd10$ICD10Code,500,replace=T),
  icd5=sample(icd10$ICD10Code,500,replace=T),
  icd6=sample(icd10$ICD10Code,500,replace=T),
  icd7=sample(icd10$ICD10Code,500,replace=T),
  icd8=sample(icd10$ICD10Code,500,replace=T),
  icd9=sample(icd10$ICD10Code,500,replace=T),
  icd10=sample(icd10$ICD10Code,500,replace=T),
  icd11=sample(icd10$ICD10Code,500,replace=T),
  icd12=sample(icd10$ICD10Code,500,replace=T),
  icd13=sample(icd10$ICD10Code,500,replace=T),
  icd14=sample(icd10$ICD10Code,500,replace=T),
  icd15=sample(icd10$ICD10Code,500,replace=T),
  icd16=sample(icd10$ICD10Code,500,replace=T))

#### Factored data ####
ind.f<-ind%>%
  # Changing coding to descriptive factors
  mutate(batch=factor(batch,levels=c(1:2)),
         male=factor(male,levels=c(0:1),labels=c("FEMALE","MALE")),
         race=factor(race,levels=c(1:7),
                     labels=c("NON-HISPANIC WHITE",
                              "NON-HISPANIC BLACK",
                              "HISPANIC / LATINO",
                              "ASIAN",
                              "NATIVE AMERICAN / AMERICAN INDIAN / ALASKA NATIVE",
                              "MULTI-RACIAL",
                              "OTHER")),
         edu=factor(edu,levels=c(1:5),
                    labels=c("LESS THAN HIGHSCHOOL",
                             "HIGH SCHOOL / GED / CERTIFICATE / SOME COLLEGE",
                             "BACHELOR'S / ASSOCIATE'S DEGREE",
                             "MASTER'S DEGREE OR HIGHER",
                             "SPECIAL EDUCATION")),
         immig=factor(immig,levels=c(0:4),
                      labels=c("BORN AND LIVE IN US",
                               "BORN ELSEWHERE AND LIVE IN US",
                               "BORN ELSEWHERE AND LIVE ELSEWHERE",
                               "BORN IN US AND LIVE ELSEWHERE",
                               "BORN ELSEWHERE")),
         pimmig=factor(pimmig,levels=c(0:2),
                              labels=c("BOTH PARENTS BORN IN US",
                                       "AT LEAST ONE PARENT BORN OUTSIDE US",
                                       "BOTH PARENTS BORN OUTSIDE US")),
         marital=factor(marital,levels=c(1,3:5),
                        labels=c("MARRIED OR SEPERATED",
                                 "WIDOWED",
                                 "DIVORCED",
                                 "NEVER MARRIED")),
         veteran=factor(veteran,levels=c(0:1),
                        labels=c("NOT A VETERAN",
                                 "VETERAN")),
         preg=factor(preg,levels=c(0:2),
                     labels=c("NOT PREGNANT IN LAST YEAR",
                              "PREGNANT AT DEATH",
                              "NOT PREGNANT AT DEATH, PREGNANT IN LAST YEAR")),
         dplace=factor(dplace,levels=c(1:8),
                       labels=c("HOSPITAL, INPATIENT",
                                "HOSPITAL, OUTPATIENT / ER",
                                "HOSPITAL, DOA",
                                "RESIDENCE",
                                "HOSPICE",
                                "NURSING HOME",
                                "ASSISTED LIVING FACILITY / REST HOME",
                                "OTHER")),
         travel=factor(travel,levels=c(0:1),
                       labels=c("DIED AND RESIDE IN SAME CITY",
                                "DIED AND RESIDE IN DIFFERENT CITIES")))

#### Class conversions ####
ind.f$dcity<-as.character(ind.f$dcity)
ind.f$ddad<-as.character(ind.f$ddad)
# ind$ddate<-as.character(ind$ddate)
ind.f$dfacilitynum<-as.character(ind.f$dfacilitynum)
ind.f$dnat<-as.character(ind.f$dnat)
ind.f$dstate<-as.character(ind.f$dstate)
ind.f$dzip<-as.character(ind.f$dzip)
ind.f$icd1<-as.character(ind.f$icd1)
ind.f$icd2<-as.character(ind.f$icd2)
ind.f$icd3<-as.character(ind.f$icd3)
ind.f$icd4<-as.character(ind.f$icd4)
ind.f$icd5<-as.character(ind.f$icd5)
ind.f$icd6<-as.character(ind.f$icd6)
ind.f$icd7<-as.character(ind.f$icd7)
ind.f$icd8<-as.character(ind.f$icd8)
ind.f$icd9<-as.character(ind.f$icd9)
ind.f$icd10<-as.character(ind.f$icd10)
ind.f$icd11<-as.character(ind.f$icd11)
ind.f$icd12<-as.character(ind.f$icd12)
ind.f$icd13<-as.character(ind.f$icd13)
ind.f$icd14<-as.character(ind.f$icd14)
ind.f$icd15<-as.character(ind.f$icd15)
ind.f$icd16<-as.character(ind.f$icd16)
ind.f$indust<-as.character(ind.f$indust)
ind.f$occup<-as.character(ind.f$occup)
ind.f$resadd<-as.character(ind.f$resadd)
ind.f$rescity<-as.character(ind.f$rescity)
ind.f$resnat<-as.character(ind.f$resnat)
ind.f$resstate<-as.character(ind.f$resstate)
ind.f$reszip<-as.character(ind.f$reszip)
ind.f$sfnum<-as.character(ind.f$sfnum)

head(ind.f)
```

\newpage

## Aggregate Data Set

```{r}
# All possible combination of month, year, and zip
base<-data.frame(zip=rep(shp@data$ZCTA5CE10,(12*18)),
                 month=rep(c(rep(1,538),rep(2,538),rep(3,538),rep(4,538),
                         rep(5,538),rep(6,538),rep(7,538),rep(8,538),
                         rep(9,538),rep(10,538),rep(11,538),rep(12,538)),18),
                 year=c(rep(2000,538*12),rep(2001,538*12),rep(2002,538*12),
                        rep(2003,538*12),rep(2004,538*12),rep(2005,538*12),
                        rep(2006,538*12),rep(2007,538*12),rep(2008,538*12),
                        rep(2009,538*12),rep(2010,538*12),rep(2011,538*12),
                        rep(2012,538*12),rep(2013,538*12),rep(2014,538*12),
                        rep(2015,538*12),rep(2016,538*12),rep(2017,538*12)))
agMonth<-base%>%
  # join to aggregated counts
  left_join(ind%>%
              # Extract death year and month from date object
              mutate(dyear=year(as.Date(ddate)),
                     dmonth=month(as.Date(ddate)))%>%
              # Determine number of cases in each zip code in each month
              group_by(reszip,dyear,dmonth)%>%
              summarize(cases=n()),
            by=c("zip"="reszip","month"="dmonth","year"="dyear"))%>%
  # Turn NA to 0
  mutate(cases=map_dbl(.x=cases,.f=~if(is.na(.x)){return(0)}else{return(.x)}))

head(agMonth)
```


\newpage

# Exploratory Data Analysis

## Summary Tables

These tables give an overview of the data based on the type of data it is. Helpful during the data cleaning stage when deciding on possible class conversions. 

```{r}
#### Numeric Variables ####
ind.f%>%
  select(names(ind[map_lgl(ind.f,is.numeric)]))%>%
  gather(colnames(ind[map_lgl(ind.f,is.numeric)]),key=variable,value=value)%>%
  group_by(variable)%>%
  summarize(Mean=mean(value,na.rm=T),
  SD=sd(value,na.rm=T),
  R1=range(value,na.rm=T)[1],
  R2=range(value,na.rm=T)[2],
  UniqueValues=length(unique(value[!is.na(value)])),
  PropMissingness=sum(is.na(value))/length(value))%>%
  mutate(Range=paste0("[",R1,", ",R2,"]"),
         percent=paste0(round(PropMissingness*100,3),"%"))%>%
  select(Variable=variable,Mean,SD,Range,`Unique Values`=UniqueValues,
         Missingness=percent)%>%
  kable(booktabs=T,digits=3,
        caption="Summary of Quantitative Variables in Individual Data Set",align="c")%>%
  kable_styling(latex_options=c("HOLD_position","striped"),position="center")

#### Character variables ####
ind.f%>%
  select(names(ind.f[map_lgl(ind.f,is.character)]))%>%
  gather(variable,value)%>%
  group_by(variable)%>%
  summarize(UniqueValues=length(unique(value[!is.na(value)])),
  PropMissingness=sum(is.na(value))/length(value))%>%
  mutate(percent=paste0(round(PropMissingness*100,3),"%"))%>%
  select(Variable=variable,`Unique Values`=UniqueValues,Missingness=percent)%>%
  kable(booktabs=T,digits=3,
        caption="Summary of Character Variables in Individual Data Set",align="c")%>%
  kable_styling(latex_options=c("HOLD_position","striped"),position="center")

#### Factor variables ####
var<-names(ind.f)[map_lgl(ind.f,is.factor)]
fac<-data.frame()
for(i in 1:11){
  df<-as.data.frame(table(ind.f[,var[i]],useNA="always"))
  df$name<-var[i]
  fac<-bind_rows(fac,df)
}
fac%>%
  select(name,everything())%>%
  mutate(Freq=prettyNum(Freq,big.mark=","))%>%
  kable(booktabs=T,longtable=T,digits=3,
        caption="Summary of Categorical Variables in Individual Data Set",
        align="c",
        col.names=c("Variable","Value","Frequency"))%>%
  kable_styling(latex_options=c("HOLD_position","repeat_header","striped"),
                position="center")%>%
  collapse_rows(columns=1,latex_hline="major",valign="middle")
```

## Univariate Graphics Loop

Plots all variables with appropriate graphical summaries. Used to find possible relationships for later investigation. Not all of these graphs are insightful, so an argument (`rmCol`) is available is specify variables for the function to ignore.

```{r}
uni.plots<-function(dataset,rmCol=NULL){
  if(is.null(rmCol)){data<-dataset}
  else{data<-dataset[,-rmCol]}
  plots<-list()
  for (i in 1:length(data)){
    if (is.numeric(data[[i]])){
      plots[[i]]<- ggplot(data,aes_string(x=colnames(data[i])))+
                    geom_histogram()+
                    xlab(colnames(data[i]))+
                    ylab("count")+
                    theme(text=element_text(family="LM Roman 10"))
                    #axis.text.x=element_text(angle=45))
    } else {
      plots[[i]]<- ggplot(data,aes_string(x=colnames(data[i])))+
      geom_bar()+
      xlab(colnames(data[i]))+
      ylab("count")+
      theme(text=element_text(family="LM Roman 10"))
            #axis.text.x=element_text(angle=45))
    }
  }
  return(marrangeGrob(plots,nrow=5,ncol=2,
                      top=textGrob("Univariate Plots",
                                   gp=gpar(fontsize=16,
                                           fontfamily="LM Roman 10"))))
}

uni.plots(ind,c(2,7:8,15:19,21:26,28:length(ind)))
```


\blandscape

## Cause of Death Tables

```{r}
#### Joining in ICD10 descriptions ####
ind.f<-ind.f%>%
  left_join(icd10[,c(2,5)],by=c("icd1"="ICD10Code"))%>%
  rename(icd1.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd2"="ICD10Code"))%>%
  rename(icd2.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd3"="ICD10Code"))%>%
  rename(icd3.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd4"="ICD10Code"))%>%
  rename(icd4.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd5"="ICD10Code"))%>%
  rename(icd5.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd6"="ICD10Code"))%>%
  rename(icd6.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd7"="ICD10Code"))%>%
  rename(icd7.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd8"="ICD10Code"))%>%
  rename(icd8.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd9"="ICD10Code"))%>%
  rename(icd9.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd10"="ICD10Code"))%>%
  rename(icd10.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd11"="ICD10Code"))%>%
  rename(icd11.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd12"="ICD10Code"))%>%
  rename(icd12.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd13"="ICD10Code"))%>%
  rename(icd13.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd14"="ICD10Code"))%>%
  rename(icd14.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd15"="ICD10Code"))%>%
  rename(icd15.desc=LongDesc)%>%
  left_join(icd10[,c(2,5)],by=c("icd16"="ICD10Code"))%>%
  rename(icd16.desc=LongDesc)

#### Primary ####
as.data.frame(table(ind.f$icd1.desc))%>%
  top_n(5,Freq)%>%
  kable(booktabs=T,digits=3,longtable=T,caption="Summary of Primary Causes of Death",align="c",
        col.names=c("Cause of Death","Frequency"))%>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),position="center")%>%
  column_spec(1,width="8in")

#### Secondary ####
as.data.frame(table(ind.f$icd2.desc))%>%
  top_n(5,Freq)%>%
  kable(booktabs=T,digits=3,longtable=T,caption="Summary of Secondary Causes of Death",align="c",
        col.names=c("Cause of Death","Frequency"))%>%
  kable_styling(latex_options=c("HOLD_position","striped","repeat_header"),position="center")%>%
  column_spec(1,width="8in")
```

\elandscape

## Mapping

```{r CountMapYear, fig.height=6}
map<-fortify(shp,region="ZCTA5CE10")
map<- merge(map, agMonth%>%rename(id=zip)%>%group_by(id,year)%>%
              summarize(cases=sum(cases)), by="id", all.x = TRUE)
map<- map[order(map$order),]
# 2000-2005
ggplot()+
  geom_polygon(data=map[map$year<=2005,],
               aes(long,lat,group=group,
                   fill=cut(cases,c(-1,0,5,10,15,20,25,30,35))),color=NA)+
  coord_fixed(1.2)+
  facet_wrap(~year,ncol=2)+
  scale_fill_viridis(discrete=T,drop=F,na.value="antiquewhite",
                     labels =c("0","(0,5]","(5,10]","(10,15]",
                               "(15,20]","(20,25]","(25,30]",
                               "(30,35]"))+
  labs(x="Longitude",y="Latitude",fill=NULL,
       title ="Observed Opioid Overdose Death Counts in MA")+
  theme(legend.position = "bottom",legend.key.width=unit(2.5,"cm"),
        legend.spacing.x = unit(.3, 'cm'),plot.title = element_text(hjust = 0.5),
        text=element_text(family="LM Roman 10"))
```

\newpage

```{r, fig.height=7}
#2006-2011
ggplot()+
  geom_polygon(data=map[map$year<=2011&map$year>=2006,],
               aes(long,lat,group=group,
                   fill=cut(cases,c(-1,0,5,10,15,20,25,30,35))),color=NA)+
  coord_fixed(1.2)+
  facet_wrap(~year,ncol=2)+
  scale_fill_viridis(discrete=T,drop=F,na.value="antiquewhite",
                     labels =c("0","(0,5]","(5,10]","(10,15]",
                               "(15,20]","(20,25]","(25,30]",
                               "(30,35]"))+
  labs(x="Longitude",y="Latitude",fill=NULL,
       title ="Observed Opioid Overdose Death Counts in MA")+
  theme(legend.position = "bottom",legend.key.width=unit(2.5,"cm"),
        legend.spacing.x = unit(.3, 'cm'),plot.title = element_text(hjust = 0.5),
        text=element_text(family="LM Roman 10"))
```

\newpage

```{r, fig.height=7}
#2012-2017
ggplot()+
  geom_polygon(data=map[map$year<=2017&map$year>=2012,],
               aes(long,lat,group=group,
                   fill=cut(cases,c(-1,0,5,10,15,20,25,30,35))),color=NA)+
  coord_fixed(1.2)+
  facet_wrap(~year,ncol=2)+
  scale_fill_viridis(discrete=T,drop=F,na.value="antiquewhite",
                     labels =c("0","(0,5]","(5,10]","(10,15]",
                               "(15,20]","(20,25]","(25,30]",
                               "(30,35]"))+
  labs(x="Longitude",y="Latitude",fill=NULL,
       title ="Observed Opioid Overdose Death Counts by Year")+
  theme(legend.position = "bottom",legend.key.width=unit(2.5,"cm"),
        legend.spacing.x = unit(.3, 'cm'),plot.title = element_text(hjust = 0.5),
        text=element_text(family="LM Roman 10"))
```

\newpage

```{r CountMapMonth, fig.height=6}
map<-fortify(shp,region="ZCTA5CE10")
map<- merge(map, 
            agMonth%>%rename(id=zip)%>%mutate(month=month(month,label=T,abbr=F))%>%
              group_by(id,month)%>%summarize(cases=sum(cases)), by="id", all.x = TRUE)
map<- map[order(map$order),]
# Jan-Jul
ggplot()+
  geom_polygon(data=map[map$month=="January"|map$month=="February"|map$month=="March"|
                        map$month=="April"|map$month=="May"|map$month=="June",],
               aes(long,lat,group=group,
                   fill=cut(cases,c(-1,0,5,10,15,20,25,30,35))),color=NA)+
  coord_fixed(1.2)+
  facet_wrap(~month,ncol=2)+
  scale_fill_viridis(discrete=T,drop=F,na.value="antiquewhite",
                     labels =c("0","(0,5]","(5,10]","(10,15]","(15,20]","(20,25]",
                               "(25,30]","(30,35]"))+
  labs(x="Longitude",y="Latitude",fill=NULL,
       title ="Observed Opioid Overdose Death Counts in MA")+
  theme(legend.position = "bottom",legend.key.width=unit(2.5,"cm"),
        legend.spacing.x = unit(.3, 'cm'),plot.title = element_text(hjust = 0.5),
        text=element_text(family="LM Roman 10"))
```

\newpage

```{r, fig.height=7}
# Aug-Dec
ggplot()+
  geom_polygon(data=map[map$month=="July"|map$month=="August"|
                        map$month=="September"|map$month=="October"|
                        map$month=="November"|map$month=="December",],
               aes(long,lat,group=group,
                   fill=cut(cases,c(-1,0,5,10,15,20,25,30,35))),color=NA)+
  coord_fixed(1.2)+
  facet_wrap(~month,ncol=2)+
  scale_fill_viridis(discrete=T,drop=F,na.value="antiquewhite",
                     labels =c("0","(0,5]","(5,10]","(10,15]","(15,20]",
                               "(20,25]","(25,30]","(30,35]"))+
  labs(x="Longitude",y="Latitude",fill=NULL,
       title ="Observed Opioid Overdose Death Counts in MA")+
  theme(legend.position = "bottom",legend.key.width=unit(2.5,"cm"),
        legend.spacing.x = unit(.3, 'cm'),plot.title = element_text(hjust = 0.5),
        text=element_text(family="LM Roman 10"))
```
